\papertitle

### 2024-12-06 src/tex/ms.tex
Fix blocking figure
Typo
Formatting
Fix references

\let\oldaddcontentsline\addcontentsline
\renewcommand{\addcontentsline}[3]{}

@@

\author{Andrew R. McCluskey}
\email{andrew.mccluskey@bristol.ac.uk}
  \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK.}
  \affiliation{European Spallation Source ERIC, Data Management and Software Centre, Asmussens Allé 305, DK-2800 Kongens Lyngby, DK.}
  \affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK.}
\author{Samuel W. Coles}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK}
  \affiliation{The Faraday Institution, Quad One, Harwell Science and Innovation Campus, Didcot, OX11 0RA, UK}
\author{Benjamin J. Morgan}
  \email{b.j.morgan@bath.ac.uk}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK}
  \affiliation{The Faraday Institution, Quad One, Harwell Science and Innovation Campus, Didcot, OX11 0RA, UK}

@@

One commonly used parameter for quantifying atomic-scale mass transport is the self-diffusion coefficient, $\D$, which describes diffusion in the absence of a chemical potential gradient.
$\D$ is related to the ensemble-average mean squared displacement (MSD), $\MSD{t}$, via the Einstein relation~\cite{einstein_uber_1905,helfand_transport_1960},
\begin{equation}
    \D = \lim_{t\to\infty}\frac{\MSD{t}}{6t},
    \label{equ:einstein}
\end{equation}
where $\Delta\mathbf{r}{(t)}$ is the displacement of a diffusion particle in the time interval $t$.
Because numerical simulations are finite in time and space, MSDs obtained from simulation data always differ from the true ensemble average MSD.
One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in \cref{equ:einstein}~\footnote{While estimating $\D$ by fitting a linear model to MSD data is, perhaps, the most commonly used approach in the literature, alternative methods that avoid linear fitting of MSD data also exist; see, for example, Refs.~{\protect\cite{VestergaardEtAl_PhysRevE2014,KrapfEtAl_NewJournalPhysics2018,bullerjohn_optimal_2020}}.}.

@@

Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while accurately estimating the corresponding statistical uncertainty using data from a single simulation.
We model the statistical population of simulation MSDs as a multivariate normal distribution, using an analytical covariance matrix derived for an equivalent system of freely diffusing particles, with this covariance matrix parameterised from the observed simulation data.
We then use Markov-chain Monte Carlo to sample the posterior distribution of linear models compatible with this multivariate normal model.
The resulting posterior distribution provides an efficient estimate for $\D$ and allows the associated statistical uncertainty in $\Dest$ to be accurately quantified.
This method is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}.

@@

The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\footnote{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}. In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:varestMSD}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories. For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}.
To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (\cref{fig:msd}a).
We also compute the MSD and estimated variance using data from a single simulation trajectory (\cref{fig:msd}b), using the scheme described above.
A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in \cref{fig:msd}c: the close numerical agreement confirms that \cref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via \cref{equ:cvv}.
\begin{figure}
    \centering
    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
    \includegraphics[width=\columnwidth]{msd.pdf}
    \caption{
        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (\cref{equ:varestMSD}).
        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via \cref{equ:varestMSD}.
        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
    }
    \label{fig:msd}
    \script{msd.py}
\end{figure}

@@

The approximate Bayesian regression scheme presented here provides a statistically efficient means of estimating the self-diffusion coefficient, $\D$, from molecular dynamics simulation data.
It improves upon textbook approaches by providing accurate point estimates of $\D$ with near-optimal statistical efficiency, while also providing a reasonable description of the uncertainty in these estimates.
The high statistical efficiency of our method allows for the use of smaller simulations, which can significantly reduce computational costs.
Overall, our method offers significant advantages over more conventional methods of estimating self-diffusion coefficients from atomistic simulations.
We have implemented this procedure in the open-source package \textsc{kinisi} \cite{mccluskey_kinisi_2022}, which we hope will support its use within the broader simulation community.

@@

Derivation of long-time limit covariance matrix for a system of freely diffusing particles, comparison of variance rescaling (\cref{equ:varestMSD}) and block renormalisation approaches, discussion of bias in the distribution of estimated variance of the estimated diffusion coefficient, and a comparison of OLS, WLS, and GLS as estimators for $\D$ applied to MSD data from simulations of \ce{Li7La3Zr2O12} (LLZO).
Additional Electronic Supplementary Information (ESI) available at Ref.~\cite{mccluskey_github_2022} under an MIT license: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}.
The LLZO raw simulation trajectories are available on Zenodo shared under a CC BY-SA 4.0 licence~\cite{coles_llzo_zenodo_2022}.
The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{mccluskey_kinisi_2022}, which is available under an MIT license.

@@

 \appendix
 \renewcommand\thesection{SI.\Roman{section}}
 \counterwithout{figure}{section}
 \renewcommand\thefigure{SI.\arabic{figure}}
 \setcounter{figure}{0}
 \counterwithout{equation}{section}
 \renewcommand\theequation{SI.\arabic{equation}}
 \setcounter{equation}{0}
 \counterwithout{table}{section}
 \renewcommand\thetable{SI.\arabic{table}}
 \setcounter{table}{0}
 \pagenumbering{arabic} 
 \renewcommand\thepage{SI.\arabic{page}}
 \addtocontents{toc}{\protect\setcounter{tocdepth}{0}}

@@

\begin{figure*}[tb]
  \centering
%   \resizebox{11cm}{!}{\includegraphics*{figure3_SI.pdf}} 
  \includegraphics[width=\columnwidth]{pyblock.pdf}
    \caption{\label{fig:msd_blocking_comparison}Comparison of the numerical variance in observed MSD from multiple replica simulations (panel a), the estimated variance obtained by rescaling the variance in observed squared displacements from a single simulation (panels b and c) (Eqn.~7), and the estimated variance obtained from the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}, as implemented in \textsc{pyblock}~\cite{spencer_pyblock_2020}. Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of 128 steps per particle, with error bars of $\pm2\var{\oMSDi}$. Panel (b) shows the MSD from just one simulation, with error bars of $\pm\varest{\oMSDi}$, obtained via Eqn.~7. Panel (d) shows the same one-simulation MSD, again with error bars of $\pm\varest{\oMSDi}$, obtained using the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}. Panels (c) and (e) plot the numerical variance against the single-simulation estimated variances obtained with each method, as a function of timestep, $i$.}
    \script{pyblock.py}
\end{figure*}

\begin{figure*}[tb]
  \centering
%   \resizebox{10.0cm}{!}{\includegraphics*{figure4_SI.pdf}} 
\includegraphics[width=\columnwidth]{msd_blocking.pdf}
    \caption{\label{fig:pyblock_comparison}(Panels (a) and (c)) Probability distributions of point-estimates $\prob{\D}$ obtained from \num{4096} individual random-walk simulations. Panel (a) shows data obtained using variance rescaling to estimated the variances, $\varest{\oMSDi}$, used to parametrise the model covariance matrix, $\mathbf{\Sigma^\prime}$. Panel (b) shows data obtained using block renormalisation to estimate the variances, $\varest{\oMSDi}$, used to parametrise the model covariance matrix.
    The grey lines show the distribution of point estimates, $\prob{\Dest}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations. 
    The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$. 
    Panels (b) and (d) show the corresponding  probability distributions of estimated variances, $\varest{\Dest}$, 
    for individual random-walk simulations, compared to the true sample variances obtained using each method (pink vertical lines) 
    $\var{\Dest}$.
    }
    \script{msd_blocking.py}
\end{figure*}

@@

\begin{figure}[tb]
  \centering
  % \resizebox{8cm}{!}{\includegraphics*{figure6_SI.pdf}} 
  \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
    \caption{Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS), from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with \num{56} lithium ions. In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\var{\Dest}$ for each regression method.}
    \label{fig:glswlsols_llzo}
    \script{glswlsols_llzo.py}
\end{figure}

>>>>

\let\oldaddcontentsline\addcontentsline
\renewcommand{\addcontentsline}[3]{}
\footnotetext[1]{While estimating $\D$ by fitting a linear model to MSD data is, perhaps, the most commonly used approach in the literature, alternative methods that avoid linear fitting of MSD data also exist; see, for example, Refs.~{\protect\cite{VestergaardEtAl_PhysRevE2014,KrapfEtAl_NewJournalPhysics2018,bullerjohn_optimal_2020}}.}
\footnotetext[2]{A common approach to estimating the variance of the mean of time-correlated data with unknown correlation time is the renormalization group blocking method of Flyvberg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}. In the SI, we compare this blocking method to our direct rescaling method (\cref{equ:varestMSD}) for estimating $\var{\oMSDi}$ from a set of random-walk trajectories. For our example data, direct rescaling performs better (gives more accurate estimates of $\var{\oMSDi}$).}

@@

\author{Andrew R. McCluskey}
\email{andrew.mccluskey@bristol.ac.uk}
  \affiliation{Centre for Computational Chemistry, School of Chemistry, University of Bristol, Cantock's Close, Bristol, BS8 1TS, UK.}
  \affiliation{European Spallation Source ERIC, Data Management and Software Centre, Asmussens Allé 305, DK-2800 Kongens Lyngby, DK.}
  \affiliation{Diamond Light Source, Harwell Campus, Didcot, OX11 0DE, UK.}
\author{Samuel W. Coles}
  \thanks{Present Address: Yusuf Hamied Department of Chemistry, University of Cambridge, Lensfield Road, Cambridge, CB2 1EW, UK.}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK}
  \affiliation{The Faraday Institution, Quad One, Harwell Science and Innovation Campus, Didcot, OX11 0RA, UK}
\author{Benjamin J. Morgan}
  \email{b.j.morgan@bath.ac.uk}
  \affiliation{Department of Chemistry, University of Bath, Claverton Down, Bath, BA2 7AY, UK}
  \affiliation{The Faraday Institution, Quad One, Harwell Science and Innovation Campus, Didcot, OX11 0RA, UK}

@@

One commonly used parameter for quantifying atomic-scale mass transport is the self-diffusion coefficient, $\D$, which describes diffusion in the absence of a chemical potential gradient.
$\D$ is related to the ensemble-average mean squared displacement (MSD), $\MSD{t}$, via the Einstein relation~\cite{einstein_uber_1905,helfand_transport_1960},
\begin{equation}
    \D = \lim_{t\to\infty}\frac{\MSD{t}}{6t},
    \label{equ:einstein}
\end{equation}
where $\Delta\mathbf{r}{(t)}$ is the displacement of a diffusing particle in the time interval $t$.
Because numerical simulations are finite in time and space, MSDs obtained from simulation data always differ from the true ensemble average MSD.
One can, however, compute an estimate of the self-diffusion coefficient, $\Dest$, by fitting a linear model to the observed MSD and using the gradient of this fitted model in place of $\MSD{t} / t$ in \cref{equ:einstein}~\cite{Note1,VestergaardEtAl_PhysRevE2014,KrapfEtAl_NewJournalPhysics2018,bullerjohn_optimal_2020}.

@@

Here, we describe an approximate Bayesian regression method for estimating $\D$ with near-maximal statistical efficiency while accurately estimating the corresponding statistical uncertainty using data from a single simulation.
We model the statistical population of simulation MSDs as a multivariate normal distribution, using an analytical covariance matrix derived for an equivalent system of freely diffusing particles, with this covariance matrix parameterised from the observed simulation data.
We then use Markov-chain Monte Carlo to sample the posterior distribution of linear models compatible with this multivariate normal model.
The resulting posterior distribution provides an efficient estimate for $\D$ and allows the associated statistical uncertainty in $\Dest$ to be accurately quantified.
This method is implemented in the open-source Python package \textsc{kinisi}~\cite{McCluskey2024}.

@@

The estimated variance $\varest{\oMSD}$ can be calculated from a single simulation trajectory, and provides an accurate estimate of the true variance $\var{\oMSD}$~\cite{Note2,FlyvbjergAndPetersen_JChemPhys1989,Frenkel2023-ah}.
To demonstrate this, we performed \num{4096} independent simulations of \num{128} particles undergoing a three-dimensional cubic-lattice random walk of \num{128} steps per particle.
Using data from all \num{4096} simulations, we first compute the true simulation MSD and its variance (\cref{fig:msd}a).
We also compute the MSD and estimated variance using data from a single simulation trajectory (\cref{fig:msd}b), using the scheme described above.
A quantitative comparison between the true MSD variance and the single-trajectory estimated MSD variance is made in \cref{fig:msd}c: the close numerical agreement confirms that \cref{equ:varestMSD} can be used to estimate $\var{\oMSD}$, which can then be used to parameterise the model covariance matrix $\mathbf{\Sigma^\prime}$ via \cref{equ:cvv}.
\begin{figure}
    \centering
    % \resizebox{\columnwidth}{!}{\includegraphics*{figure2.pdf}}
    \includegraphics[width=\columnwidth]{msd.pdf}
    \caption{
        Comparison of the numerical variance in observed MSD from multiple replica simulations and the estimated variance in observed MSD given by rescaling the variance in observed squared displacements (\cref{equ:varestMSD}).
        Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of \num{128} steps per particle, with error bars of $\pm2\sigma[\oMSDi]$.
        Panel (b) shows the MSD from just one simulation, with error bars of $\pm2\widehat{\sigma}[\oMSDi]$, obtained via \cref{equ:varestMSD}.
        Panel (c) plots the numerical variance against the estimated variance from a single simulation as a function of timestep $i$.
    }
    \label{fig:msd}
    \script{msd.py}
\end{figure}

@@

The approximate Bayesian regression scheme presented here provides a statistically efficient means of estimating the self-diffusion coefficient, $\D$, from molecular dynamics simulation data.
It improves upon textbook approaches by providing accurate point estimates of $\D$ with near-optimal statistical efficiency, while also providing a reasonable description of the uncertainty in these estimates.
The high statistical efficiency of our method allows for the use of smaller simulations, which can significantly reduce computational costs.
Overall, our method offers significant advantages over more conventional methods of estimating self-diffusion coefficients from atomistic simulations.
We have implemented this procedure in the open-source package \textsc{kinisi} \cite{McCluskey2024}, which we hope will support its use within the broader simulation community.

@@

Derivation of long-time limit covariance matrix for a system of freely diffusing particles, comparison of variance rescaling (\cref{equ:varestMSD}) and block renormalisation approaches, discussion of bias in the distribution of estimated variance of the estimated diffusion coefficient, and a comparison of OLS, WLS, and GLS as estimators for $\D$ applied to MSD data from simulations of \ce{Li7La3Zr2O12} (LLZO).
Additional Electronic Supplementary Information (ESI) available at Ref.~\cite{mccluskey_github_2022} under an MIT license: A complete set of analysis/plotting scripts allowing for a fully reproducible and automated analysis workflow, using \textsc{showyourwork}~\cite{luger_showyourwork_2021}.
The LLZO raw simulation trajectories are available on Zenodo shared under a CC BY-SA 4.0 licence~\cite{coles_llzo_zenodo_2022}.
The method outlined in this work is implemented in the open-source Python package \textsc{kinisi}~\cite{McCluskey2024}, which is available under an MIT license.

@@

 \appendix
 \renewcommand\thesection{S-\Roman{section}}
 \counterwithout{figure}{section}
 \renewcommand\thefigure{S-\arabic{figure}}
 \setcounter{figure}{0}
 \counterwithout{equation}{section}
 \renewcommand\theequation{S-\arabic{equation}}
 \setcounter{equation}{0}
 \counterwithout{table}{section}
 \renewcommand\thetable{S-\arabic{table}}
 \setcounter{table}{0}
 \pagenumbering{arabic} 
 \renewcommand\thepage{S-\arabic{page}}
 \addtocontents{toc}{\protect\setcounter{tocdepth}{0}}

@@

\begin{figure*}[tb]
  \centering
%   \resizebox{11cm}{!}{\includegraphics*{figure3_SI.pdf}} 
  \includegraphics[width=\textwidth]{msd_blocking.pdf}
    \caption{\label{fig:msd_blocking_comparison}Comparison of the numerical variance in observed MSD from multiple replica simulations (panel a), the estimated variance obtained by rescaling the variance in observed squared displacements from a single simulation (panels b and c) (Eqn.~7), and the estimated variance obtained from the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}, as implemented in \textsc{pyblock}~\cite{spencer_pyblock_2020}. Panel (a) shows the mean observed MSD from \num{4096} simulations of \num{128} particles undergoing a 3D lattice random walk of 128 steps per particle, with error bars of $\pm2\var{\oMSDi}$. Panel (b) shows the MSD from just one simulation, with error bars of $\pm\varest{\oMSDi}$, obtained via Eqn.~7. Panel (d) shows the same one-simulation MSD, again with error bars of $\pm\varest{\oMSDi}$, obtained using the block renormalisation method of Flyvbjerg and Peterson~\cite{FlyvbjergAndPetersen_JChemPhys1989}. Panels (c) and (e) plot the numerical variance against the single-simulation estimated variances obtained with each method, as a function of timestep, $i$.}
    \script{msd_blocking.py}
\end{figure*}

\begin{figure*}[tb]
  \centering
%   \resizebox{10.0cm}{!}{\includegraphics*{figure4_SI.pdf}} 
\includegraphics[width=\columnwidth]{pyblock.pdf}
    \caption{\label{fig:pyblock_comparison}(Panels (a) and (c)) Probability distributions of point-estimates $\prob{\D}$ obtained from \num{4096} individual random-walk simulations. Panel (a) shows data obtained using variance rescaling to estimated the variances, $\varest{\oMSDi}$, used to parametrise the model covariance matrix, $\mathbf{\Sigma^\prime}$. Panel (b) shows data obtained using block renormalisation to estimate the variances, $\varest{\oMSDi}$, used to parametrise the model covariance matrix.
    The grey lines show the distribution of point estimates, $\prob{\Dest}$, obtained using Bayesian regression with a mean vector and numerical covariance matrix derived from the complete dataset of all \num{4096} simulations. 
    The pink horizontal bar shows an interval of one standard deviation in $\prob{\Dest}$. 
    Panels (b) and (d) show the corresponding  probability distributions of estimated variances, $\varest{\Dest}$, 
    for individual random-walk simulations, compared to the true sample variances obtained using each method (pink vertical lines) 
    $\var{\Dest}$.
    }
    \script{pyblock.py}
\end{figure*}

@@

\begin{figure}[tb]
  \centering
  % \resizebox{8cm}{!}{\includegra?phics*{figure6_SI.pdf}} 
  \includegraphics[width=\columnwidth]{glswlsols_llzo.pdf}
    \caption{Example distributions of estimated self-diffusion coefficients, $\Dest$, calculated using (a) ordinary least squares (OLS), (b) weighted least squares (WLS), and (c) generalised least squares (GLS), from MSD data from \num{512} effective simulations of LLZO of \SI{\sim 25}{\pico\second} with \num{56} lithium ions. In each panel, the grey curve shows the best-fit normal distribution for the simulation data, the upper horizontal bar shows the standard deviation of this distribution, and the lower horizontal bar shows the average estimated standard distribution given by the analytical expression for $\var{\Dest}$ for each regression method.}
    \label{fig:glswlsols_llzo}
    \script{glswlsols_llzo.py}
\end{figure}
